MHF SCP Experiment Visualisations

### 2024-09-02 minos/outcomes/MHF_SCP_visualisations.Rmd
Added plot for mental health risk with intervals based on +- 1 for threshold (i.e. 46.6 instad of 45.6). Also fixed the lone parent household plot

  p2 <- ggplot(plot_data2_confint, aes(x = year, y = mean_SF12_to_cost_ratio, group = scen)) +
    geom_ribbon(aes(ymin = lower, ymax = upper, color = scen, fill = scen), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = scen), linetype = 'dashed') +
    geom_point(aes(color = scen)) +
    scale_color_manual(values = palette) +
    scale_fill_manual(values = palette) +
    xlab('Year') +
    ylab('Mean SF-12 Change') +
    theme(
        legend.title = element_blank(),  # Remove the legend title
        legend.position = c(0.2, 0.7),  # Position the legend inside the plot
        legend.text = element_text(size = 8),  # Increase the text size for better readability
        legend.key.size = unit(0.8, 'cm'),  # Reduce the key size to fit more entries
        legend.spacing.y = unit(0.1, 'cm')  # Add vertical space between legend items
    ) +
    guides(
        color = guide_legend(
            ncol = ifelse(length(unique(plot_data1_confint$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        ),
        fill = guide_legend(
            ncol = ifelse(length(unique(plot_data1_confint$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        )
    )

@@

  lone_parent.change <- UC.lone_parent %>%
    select(run_id, sex, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    select(run_id, sex, year, diff_SF_12)

  lone_parent.confint <- lone_parent.change %>%
    group_by(sex, year) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

@@

  p1 <- ggplot(data = lone_parent.confint, aes(x = year, y = mean_SF12_change, group = sex, colour = sex, fill = sex)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = sex), linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
    scale_color_manual(values = c("#E69F00", "#56B4E9")) +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('SF-12 Change')

  p2 <- ggplot(data = lone_parent.confint, aes(x = year, y = mean_SF12_change, group = sex, colour = sex, fill = sex)) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('SF-12 Change')

@@

@@

```{r}
mental_illness_risk_plots('UC_men_illness_risk', scen.list = scen.list.25_100)
```

@@

```{r}
mental_illness_risk_plots('UC_families_men_illness_risk', scen.list = scen.list.25_100)
```

@@

>>>>

  p2 <- ggplot(plot_data2_confint, aes(x = year, y = mean_SF12_to_cost_ratio, group = scen)) +
    geom_ribbon(aes(ymin = lower, ymax = upper, color = scen, fill = scen), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = scen), linetype = 'dashed') +
    geom_point(aes(color = scen)) +
    scale_color_manual(values = palette) +
    scale_fill_manual(values = palette) +
    xlab('Year') +
    ylab('Cost - SF12 Change Ratio') +
    theme(
        legend.title = element_blank(),  # Remove the legend title
        legend.position = c(0.2, 0.7),  # Position the legend inside the plot
        legend.text = element_text(size = 8),  # Increase the text size for better readability
        legend.key.size = unit(0.8, 'cm'),  # Reduce the key size to fit more entries
        legend.spacing.y = unit(0.1, 'cm')  # Add vertical space between legend items
    ) +
    guides(
        color = guide_legend(
            ncol = ifelse(length(unique(plot_data1_confint$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        ),
        fill = guide_legend(
            ncol = ifelse(length(unique(plot_data1_confint$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        )
    )

@@

  lone_parent.change <- UC.lone_parent %>%
    select(run_id, lone_parent_flag, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    select(run_id, lone_parent_flag, year, diff_SF_12)

  lone_parent.confint <- lone_parent.change %>%
    group_by(lone_parent_flag, year) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

@@

  lone_parent.confint$lone_parent_flag <- ifelse(lone_parent.confint$lone_parent_flag == 1, 'TRUE', 'FALSE')

  p1 <- ggplot(data = lone_parent.confint, aes(x = year, y = mean_SF12_change, group = lone_parent_flag, colour = lone_parent_flag, fill = lone_parent_flag)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = lone_parent_flag), linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    scale_fill_manual(values = c("#49a04d", "#3f3d9e")) +
    scale_color_manual(values = c("#49a04d", "#3f3d9e")) +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('SF-12 Change') +
    theme(
        legend.position = c(0.2, 0.7),  # Position the legend inside the plot
        legend.text = element_text(size = 8),  # Increase the text size for better readability
        legend.key.size = unit(0.8, 'cm'),  # Reduce the key size to fit more entries
        legend.spacing.y = unit(0.1, 'cm')  # Add vertical space between legend items
    ) +
    guides(fill = guide_legend(title = 'Lone Parent Household'),
           color = guide_legend(title = 'Lone Parent Household'),
           group = guide_legend(title = 'Lone Parent Household'))

  p2 <- ggplot(data = lone_parent.confint, aes(x = year, y = mean_SF12_change, group = lone_parent_flag, colour = lone_parent_flag, fill = lone_parent_flag)) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('SF-12 Change')

@@

mental_illness_risk_plots_interval <- function(summary_group, scen.list, palette = my_palette) {
  tit <- ' '
  if (summary_group == 'UC_men_illness_risk') { tit <- 'UC Population' }
  if (summary_group == 'UC_families_men_illness_risk') { tit <- 'UC Families' }
  if (summary_group == 'men_illness_risk') { tit <- 'Whole Population' }
  if (summary_group == 'men_illness_risk_families') { tit <- 'Families' }

  # Get data for the standard threshold (45.6)
  data.mid <- do.call(rbind, generate_df_list(summary_group, scen.list))

  # # get data for the upper and lower threshold
  sum_high <- paste0(summary_group, '_higher')
  sum_low <- paste0(summary_group, '_lower')
  data.high <- do.call(rbind, generate_df_list(sum_high, scen.list))
  data.low <- do.call(rbind, generate_df_list(sum_low, scen.list))

  get_counts <- function(data) {

    plot_data <- data %>%
      filter(mental_health_risk == 'TRUE') %>%
      select(-prop, -mental_health_risk) %>%
      pivot_wider(names_from = scen,
                  values_from = count) %>%
      mutate(across(starts_with("£"), ~ . - baseline, .names = "diff_{.col}")) %>%
      select(run_id, year, starts_with('diff')) %>%
      pivot_longer(cols = starts_with("diff"),
                   names_sep = '_',
                   names_to = c("diff", "scen"),
                   values_to = 'Count') %>%
      select(-diff) %>%
      group_by(year, scen) %>%
      summarise(Count = mean(Count))

    # Extract numeric part from 'scen' column then...
    scen_numeric <- as.numeric(gsub("[^0-9]", "", plot_data$scen))
    # ...Reorder 'scen' column based on numeric values for plotting
    plot_data$scen <- factor(plot_data$scen, levels = unique(plot_data$scen[order(scen_numeric)]))

    return(plot_data)
  }

  mid <- get_counts(data.mid)
  high <- get_counts(data.high)
  low <- get_counts(data.low)

  high <- high %>%
    rename(High = Count)
  low <- low %>%
    rename(Low = Count)

  plot_data_counts <- merge(mid, high, by = c('year', 'scen'))
  plot_data_counts <- merge(plot_data_counts, low, by = c('year', 'scen'))

  rm(mid, high, low)

  p1 <- ggplot(plot_data_counts, aes(x = year, y = Count, group = scen)) +
    geom_ribbon(aes(ymin = High, ymax = Low, color = scen, fill = scen), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = scen), linetype = 'dashed') +
    geom_point(aes(color = scen)) +
    scale_color_manual(values = palette) +
    scale_fill_manual(values = palette) +
    labs(subtitle = tit) +
    xlab('Year') +
    ylab('Mean Change (Count)') +
    theme(
        legend.title = element_blank(),  # Remove the legend title
        legend.position = c(0.2, 0.7),  # Position the legend inside the plot
        legend.text = element_text(size = 8),  # Increase the text size for better readability
        legend.key.size = unit(0.8, 'cm'),  # Reduce the key size to fit more entries
        legend.spacing.y = unit(0.1, 'cm')  # Add vertical space between legend items
    ) +
    guides(
        color = guide_legend(
            ncol = ifelse(length(unique(plot_data_counts$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        ),
        fill = guide_legend(
            ncol = ifelse(length(unique(plot_data_counts$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        )
    )

  get_props <- function(data) {

    plot_data <- data %>%
      filter(mental_health_risk == 'TRUE') %>%
      select(-mental_health_risk, -count) %>%
      pivot_wider(names_from = scen,
                  values_from = prop) %>%
      mutate(across(starts_with("£"), ~ . - baseline, .names = "diff_{.col}")) %>%
      select(run_id, year, starts_with('diff')) %>%
      pivot_longer(cols = starts_with("diff"),
                   names_sep = '_',
                   names_to = c("diff", "scen"),
                   values_to = 'Proportion') %>%
      select(-diff) %>%
      group_by(year, scen) %>%
      summarise(Percentage = mean(Proportion) * 100)

    # Extract numeric part from 'scen' column then...
    scen_numeric <- as.numeric(gsub("[^0-9]", "", plot_data$scen))
    # ...Reorder 'scen' column based on numeric values for plotting
    plot_data$scen <- factor(plot_data$scen, levels = unique(plot_data$scen[order(scen_numeric)]))

    return(plot_data)
  }

  mid <- get_props(data.mid)
  high <- get_props(data.high)
  low <- get_props(data.low)

  high <- high %>%
    rename(High = Percentage)
  low <- low %>%
    rename(Low = Percentage)

  plot_data_prop <- merge(mid, high, by = c('year', 'scen'))
  plot_data_prop <- merge(plot_data_prop, low, by = c('year', 'scen'))

  rm(mid, high, low)

  p2 <- ggplot(plot_data_prop, aes(x = year, y = Percentage, group = scen)) +
    geom_ribbon(aes(ymin = High, ymax = Low, color = scen, fill = scen), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = scen), linetype = 'dashed') +
    geom_point(aes(color = scen)) +
    scale_color_manual(values = palette) +
    scale_fill_manual(values = palette) +
    labs(subtitle = tit) +
    xlab('Year') +
    ylab('Mean Change (%)') +
    theme(
        legend.title = element_blank(),  # Remove the legend title
        legend.position = c(0.2, 0.7),  # Position the legend inside the plot
        legend.text = element_text(size = 8),  # Increase the text size for better readability
        legend.key.size = unit(0.8, 'cm'),  # Reduce the key size to fit more entries
        legend.spacing.y = unit(0.1, 'cm')  # Add vertical space between legend items
    ) +
    guides(
        color = guide_legend(
            ncol = ifelse(length(unique(plot_data_prop$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        ),
        fill = guide_legend(
            ncol = ifelse(length(unique(plot_data_prop$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        )
    )

@@

```{r}
mental_illness_risk_plots('UC_men_illness_risk', scen.list = scen.list.25_100)
mental_illness_risk_plots_interval('UC_men_illness_risk', scen.list = scen.list.25_100)
```

@@

```{r}
mental_illness_risk_plots('UC_families_men_illness_risk', scen.list = scen.list.25_100)
mental_illness_risk_plots_interval('UC_families_men_illness_risk', scen.list = scen.list.25_100)
```

@@

```{r}
summary_group <- 'UC_families_men_illness_risk'
scen.list <- scen.list.25_100
palette <- my_palette

mental_illness_risk_plots_interval(summary_group, scen.list, palette)
```
